### 2016-07-10 manuscript.tex
20160710

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{color}

\spnewtheorem{assumption}{Assumption}{\bf}{\it}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}

\title{Adaptive Hierarchical Sliding Mode Control with Input Saturation for Attitude Regulation of Multi-satellite Tethered System%\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.}
}
%\subtitle{Do you have a subtitle?\\ If so, write it here}

%\titlerunning{Short form of title}        % if too long for running head

\author{Zhiqiang Ma         \and
        Guanghui Sun %etc.
}

%\authorrunning{Short form of author list} % if too long for running head

\institute{Zhiqiang Ma \at
              Research Institute of Intelligent Control and Systems, Harbin Institute of Technology \\
              \email{zhiqiangmahit@gmail.com}
                        %  \\
%             \emph{Present address:} of F. Author  %  if needed
           \and
           Guanghui Sun \at
              Research Institute of Intelligent Control and Systems, Harbin Institute of Technology
              \email{guanghuisun@hit.edu.cn}
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
This paper proposes  a  novel adaptive \begin{color}{red}hierarchical\end{color}  sliding mode control for the attitude regulation of the multi-satellite inline tethered system, where the input saturation is taken into account. The governing equations for the attitude dynamics of the three-satellite inline tethered system are derived firstly by utilizing Lagrangian mechanics theory. Considering the fact that the attitude of the central satellite can be adjusted by using the simple exponential stabilization scheme, the decoupling of the central satellite and the terminal ones is presented, and in addition, the new adaptive sliding mode control law is applied to stabilize the attitude dynamics of the two terminal satellites based on the synchronization and partial contraction theory. In the adaptive hierarchical sliding mode control design, the input is modeled as saturated input due to the fact that the flywheel torque is bounded, and meanwhile, an adaptive update rate is introduced to eliminate the effect of the saturated input and the external perturbation. The proposed control scheme can be applied on the two-satellite system to achieve fixed-point rotation. Numerical results validate the effectiveness of the proposed method.
\keywords{Multi-satellite tethered system \and Adaptive Sliding mode \and Input saturation \and Satellite attitude adjusting \and Underactuated system stabilization}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\section{Introduction}
\begin{color}{red}The tethered satellite system (TSS) consists of in-orbit satellites, which are connected by long thin space tethers. Since the conception of the TSS was addressed in 1960s, many space institutes have implemented a significant number of orbital and suborbital experiments, such as TSS-1R, SEDS, MAST and YES2~\cite{williams2012review}, to verify the feasibility of the existing studies from 1970s to 1980s. \end{color}Among other things, the TSS has been ideally qualified for various missions in the space environment, like the electrodynamic propulsion application~\cite{lanoix2005effect}, deorbiting defunct satellites~\cite{khan2014analysis}, the space tug~\cite{wen2016constrained} and the space elevator~\cite{kojima2015mission}.\par
In recent years, many investigations have been carried out about the dynamics of the TSS, providing some effective solutions to the attitude stabilization and flight formations~\cite{misra2002equilibrium,yan2012feedback,xu2012nonlinear,zhu2015dynamic,hallaj2015tethered,alary2015dynamics,wang2015coordinated,ting2016reconfiguration,cicci2016preliminary}. \begin{color}{red}Typically, for the attitude stability, the steerable thruster-based control is utilized extensively due to its convenience of equipping ~\cite{aslanov2013dynamics,jasper2014input,jasper2014tethered,aslanov2014dynamics}. However, this strategy is in a dilemma about that the work of actuators is dependent on the availability and consumption of the fuel directly~\cite{chung2008propellant1}, and moreover, in thruster-based control schemes,  the propulsive working medium ejected by thrusters has an impact on the measurement of sensors in the vicinity~\cite{hallaj2015tethered}.\end{color}\par
To alleviate these concerns mentioned above, some alternative non-pure thruster-based schemes for formation flights of the TSS have been proposed. Miller addressed the concept of the electromagnetic formation flying (EMFF), in which the high temperature superconducting technology was employed to regulate the attitude indirectly~\cite{miller2010control}. Some researches followed to achieve EMFF by using different control schemes~\cite{zhang2014adaptive,huang2015lmi}. \begin{color}{red}Zhong et al. reported the application of electrodynamic tether for nanosatellite fast deorbit in low earth orbit~\cite{zhong2013dynamics}, and also presented a piecewise two-phased optimal control scheme to achieve open-loop control trajectory optimization~\cite{zhong2014optimal}.\end{color} Yang investigated the three-inline satellite array, in which the participating satellites were equipped with the reaction wheels and the thrusters, and hence the linear force and the flywheel torque cooperated to manipulate the spacecraft~\cite{yang2015decentralized}. Natarajan and Schaub combined the Coulomb forces with conventional thrust forces to obtain a control scheme, in which the Coulomb forces predominated, and investigated the stability of the two-craft formation about different equilibrium configurations~\cite{natarajan2009hybrid}. In the view of canceling the drawback of the thrusters in the attitude regulation control, as a inter-force actuator, the utilization of flywheels has been focused on~\cite{pizarro2008dynamics}. \begin{color}{red}Compared to the thruster-based scheme, the solar sails provide the power for the flywheels to achieve a long-term control, and simultaneously the contamination to the sensitive sensors does not occur. However, due to the absence of the thrusters forces, the pure flywheel torque control causes the attitude dynamics underactuated adversely~\cite{chung2008propellant1}.\end{color}\par
Besides actuator selection, the effective control scheme is also an important component of the TSS. In recent years, many researchers strived to investigate the motion dynamics of the multi-satellite inline formation, and there have existed a variety of sophisticated linear control schemes. Chung et al. exploited geometric symmetries to derive the reduction dynamics, and addressed the decentralized linear control method for the flight tailored toward the SPECS mission~\cite{chung2007nonlinear}. They also developed a gain-scheduling control law to adjust the attitude of the multiple-satellite system without any thruster forces~\cite{chung2008propellant1}. Yang et al. proposed the decentralized fault tolerance control scheme for the attitude of the satellites in the three-inline array around the equilibrium point~\cite{yang2015decentralized}. In the aspect of nonlinear control, Chung et al. presented a backstepping control design for the multi-satellite system after reducing the model into the strict-feedback cascade normal form~\cite{chung2008propellant2}. Zhao et al. applied gauss pseudospectral method to multi-objective trajectory optimization for two spacecrafts~\cite{zhao2014gauss}. Characteristically, among the strategies mentioned above, all the inputs are referred to as unsaturated behaviors, however, in the realistic application the input saturation frequently does occur. \begin{color}{red}Even further, this unexpected situation often deteriorates the performance of regulation objective practically~\cite{zavala2007natural}. The saturation usually stems from the physical mechanics and the energy dissipation, and leads to the error between the actual output and the command signal. Hence, this case can be regarded as a potential source of the corruption~\cite{Hu2009Robust}, and should be taken in account when the controller is designed.\end{color} \begin{color}{red}In order to resist the effect of the saturation, Bo{\v{s}}kovi{\'{c}} et al. proposed global stable control algorithm with variable structure philosophy to overcome input saturation~\cite{bovskovic2001robust};\end{color} Tsiotras and Luo decoupled the general motion dynamics into two rotations, and designed a feedback control scheme to regulate the system to track a specified state in the inertial space~\cite{tsiotras2000control}; Grimm et al.~\cite{grimm2003antiwindup}, Xu and Lam~\cite{xu2001robust} paid more attention to the linear systems with the bounded input. Admittedly, introducing the saturated inputs into the control systems is complicate but meaningful for both the theoretical studies and the engineering applications.\par
In this paper, a novel adaptive sliding mode control scheme is proposed for the three-satellite inline array and the two-satellite system. This work takes into consideration the characteristics of the coupling input saturation, whereas the linear force is absent in the attitude regulation of the satellite. The remainder of the paper is organized as follows. In Section~\ref{sec:mm}, the mathematical model for the flight of three-satellite inline array is presented, and a novel MHSM scheme for the stability of the attitude dynamics is proposed, with emphasis on the solution of the input saturation in Section~\ref{sec:SMC}. In Section~\ref{sec:SMC2}, the attitude dynamics of two-satellite system is investigated, \begin{color}{red}and the stability analysis of the MHSM scheme for the dynamics is exhibited.\end{color} Finally in Section~\ref{sec:sm}, the numerical simulation is illustrated to validate the proposed control design.
\section{Mathematical Model For Three-satellite Inline Array}\label{sec:mm}
\begin{figure}
\centering
\includegraphics[width=0.75\textwidth]{paper2_Fig1.eps}
\caption{Schematic representation of three-satellite inline array}
\label{Figure:3satellite}
\end{figure}
%figure1
\begin{color}{red} The structure of the three-satellite inline array stems from NASA's Submillimeter Probe of the Evolution of Cosmic Structure (SPECS) mission, which is famous as a tethered spacecraft interferometer, and the array keeps rotating while detecting submillimeter-wave-length light from the early universe~\cite{lorenzini2006far}. The structure of the three-satellite array can be seen in Figure~\ref{Figure:3satellite}. Compared with traditional two-body tethered satellite system, the two end bodies in the array can carry same instruments to observe the information to ensure accuracy of data. Also, due to the existence of the centripetal force, it is more convenient to deploy the tether while rotating the array in the deep space. In addition, some long-term spatial gravitational experiments can be developed based on this structure.\end{color}\par
The simplified Euler's rotational equations of the motion dynamics of the three-satellite inline array follows
\begin{align}
\begin{bmatrix}
J_1 &0  &0\\
0   &J_2&0\\
0   &0  &J_3
\end{bmatrix}\begin{pmatrix}
\dot{\omega}_1\\
\dot{\omega}_2\\
\dot{\omega}_3
\end{pmatrix}+\begin{bmatrix}
0   &-\omega_3  &\omega_2\\
\omega_3   &0   &-\omega_1\\
-\omega_2   &\omega_1  &0
\end{bmatrix}\begin{bmatrix}
J_1 &0  &0\\
0   &J_2&0\\
0   &0  &J_3
\end{bmatrix}\begin{pmatrix}
{\omega}_1\\
{\omega}_2\\
{\omega}_3
\end{pmatrix}=\begin{pmatrix}
\tau_1\\
\tau_2\\
\tau_3
\end{pmatrix}\label{eq:3body}
\end{align}
where $J_1$ denotes the moment of inertia of the satellite array for the vertical axis; $J_1=3I_1+2ml^2$ in which $I_1$ refers to the moment of inertia of an individual participant in the three-body inline frame; $l$ identifies the tether length; $J_2$ has the similar expression $J_2=3I_2+2ml^2$ for the second axis. In Figure~\ref{Figure:3satellite}, the first axis is perpendicular to the rotation plane which is determined by the other two axes. Attach the third axis to the tether, and the rotation about the second axis generates the out-of-plane motion. Without loss of generality, it is quite reasonable to assume that the individual satellite is axis-symmetric thoroughly. The participating satellites arranging in a line neatly gives rise to $I_1=I_2$, and in addition the array is axis-symmetric about the third axis~\cite{chung2008propellant1}. Hence, Eq.~(\ref{eq:3body}) reduces to
\begin{align}
J_1\dot{\omega}_1-(J_1-J_3)\omega_3\omega_2=\tau_1\\
J_2\dot{\omega}_2-(J_2-J_3)\omega_3\omega_1=\tau_2\\
J_3\dot{\omega}_3=\tau_3\label{eq:3model}
\end{align}
\begin{color}{red}
which brings out that the three-satellite inline tethered system torsional rotation dynamics is irrelevant to the other two axes~\cite{chung2008propellant1}. Since the dynamic state $\omega_3$ is independent from the two axes, a simple linear feedback control can be applied to control the dynamics of the third axis, such that both $\omega_3=0$ and $\dot\omega_3 = 0$. Hence Eq.~(\ref{eq:3model}) reduces to
\end{color}
\begin{align}
\begin{split}
J_1\dot\omega_1&=\tau_1\\
J_2\dot\omega_2&=\tau_2.
\end{split}
\end{align} \begin{color}{red}From equations above, $\omega_2$ is independent from $\omega_1$. Similar to the dynamic state $\omega_3$, dynamic state $\omega_2$ can be controlled separately without affecting the in-plane dynamics~\cite{chung2007nonlinear,chung2007nonlinear1,chung2007nonlinear3}. For each satellite, three flywheels are mounted orthogonally to regulate the attitude. The flywheel which is mounted perpendicularly to the tether, governs the dynamic state $\omega_3$, and different from the other axes, the varying $\omega_3$ will not lead to the situation described in Figure~\ref{Figure:model3body}, hence a simple linear feedback control can achieve its attitude regulation. When we attention on the in-plane dynamics, the other two flywheel torques and thruster firing handle the out-plane dynamics, which can guarantee that the in-plane motion occurs without other disturbance from the other axes.\end{color} On the premise of the assumption that the out-of-plane plant is stable independently, the in-plane attitude dynamics, accordingly, can be described as a model consisting of three flywheels assembly and two tethers, as seen in Figure~\ref{Figure:model3body}. \begin{color}{red}The mathematical model expressed by using Euler-Lagrange equations is established, and the origin of the array is the satellite's geometrical center as well the mass center.\end{color} In the 2D plane, the three-body satellite array arranges along $X$-axis direction, and the positive directions of $\theta_i$ and $\phi_i$ are anticlockwise. $F_i$ denotes the $i$th thruster's linear force, and $u_i$ is defined to be the $i$th flywheel torque acting on the center of mass. $\Psi$ identifies the rotation angle, then the dynamic equation follows~\cite{chung2007nonlinear1}
\begin{align}
    &\begin{bmatrix}
        M_{11}      &{M_2^T(\theta_1,\phi_1)}      &M_2^T(\theta_2,\phi_2)\\
        M_2(\theta_1,\phi_1)      &{M_1(\phi_1)}   &O_{2\times2}\\
        M_2(\theta_2,\phi_2)      &O_{2\times2}    &M_1(\phi_2)
    \end{bmatrix}
\begin{pmatrix}
\ddot{\Psi}\\
\ddot{\theta}_1\\
\ddot{\phi}_1\\
\ddot{\theta}_2\\
\ddot{\phi}_2
\end{pmatrix}\notag\\
+
    &\begin{bmatrix}
        0           &C_3^T(\phi_1,\theta_1,\dot{\phi}_1,\dot{\theta}_1)     &C_3^T(\phi_2,\theta_2,\dot{\phi}_2,\dot{\theta}_2)     \\
        C_4(\theta_1,\phi_1)      &C_1(\phi_1,\dot{\theta}_1,\dot{\phi}_1)   &O_{2\times2}\\
        C_4(\theta_2,\phi_2)      &O_{2\times2}          &C_2(\phi_2,\dot{\theta}_2,\dot{\phi}_2)
    \end{bmatrix}
\begin{pmatrix}
\dot{\Psi}\\
\dot{\theta}_1\\
\dot{\phi}_1\\
\dot{\theta}_2\\
\dot{\phi}_2
\end{pmatrix}
=
\begin{pmatrix}
u_0\\
\tau_{\theta1}\\
\tau_{\phi1}\\
\tau_{\theta2}\\
\tau_{\phi2}
\end{pmatrix}\label{eq:3bodymodel}
\end{align}
where
\begin{align}
&M_1(\phi)
=\begin{bmatrix}
m_{11}(\phi) &m_{12}(\phi)\\
m_{21}(\phi) &m_{22}
\end{bmatrix}
=\begin{bmatrix}
I_r+ml^2+2mrl\cos{\phi}  &I_r+mrl\cos{\phi}\\
I_r+mrl\cos{\phi}       &I_r
\end{bmatrix}\label{eq:M1}\\
&C_1(\phi,\dot{\theta},\dot{\phi})=
\begin{bmatrix}
c_{11}(\phi,\dot{\phi}) &c_{12}(\phi,\dot{\phi},\dot{\theta})\\
c_{21}(\phi,\dot{\phi}) &c_{22}
\end{bmatrix}=
\begin{bmatrix}
-mrl\sin{\phi}\dot{\phi}   &-mrl\sin{\phi}(\dot{\theta}+\dot{\phi})\\
mrl\sin{\phi}\dot{\phi}    &0
\end{bmatrix}\label{eq:C1}\\
&\begin{pmatrix}
\tau_\theta\\
\tau_\phi
\end{pmatrix}
=
\begin{bmatrix}
r+l\cos{\phi}   &1\\
r               &1
\end{bmatrix}
\begin{pmatrix}
F\\
u
\end{pmatrix}\\
&M_2(\theta,\phi)   =
\begin{bmatrix}
mrl\cos{(\theta-\Psi)}+mr^2\cos{(\theta+\phi-\Psi)}\\
mr^2\cos{(\theta+\phi-\Psi)}
\end{bmatrix}\notag\\
%&M_{12}=mrl\cos{(\theta_1-\Psi)}+mr^2\cos{(\theta_1+\phi_1-\Psi)}\notag\\
%&M_{13}=mr^2\cos{(\theta_1+\phi_1-\Psi)}\notag\\
%&M_{14}=mrl\cos{(\theta_2-\Psi)}+mr^2\cos{(\theta_2+\phi_2-\Psi)}\notag\\
%&M_{15}=mr^2\cos{(\theta_2+\phi_2-\Psi)}\notag\\
&C_3(\phi,\theta,\dot{\phi},\dot{\theta})   =
\begin{bmatrix}
-mrl\sin{(\theta-\Psi)}\dot\theta-mr^2\sin{(\theta+\phi-\Psi)}(\dot\theta+\dot\phi)\\
-mr^2\sin{(\theta+\phi-\Psi)}(\dot\theta+\dot\phi)
\end{bmatrix}\notag\\
%&C_{12}=-mrl\sin{(\theta_1-\Psi)}\dot\theta_1-mr^2\sin{(\theta_1+\phi_1-\Psi)}(\dot\theta_1+\dot\phi_1)\notag\\
%&C_{13}=-mr^2\sin{(\theta_1+\phi_1-\Psi)}(\dot\theta_1+\dot\phi_1)\notag\\
%&C_{14}=-mrl\sin{(\theta_2-\Psi)}\dot\theta_2-mr^2\sin{(\theta_2+\phi_2-\Psi)}(\dot\theta_2+\dot\phi_2)\notag\\
%&C_{15}=-mr^2\sin{(\theta_2+\phi_2-\Psi)}(\dot\theta_2+\dot\phi_2)\notag\\
&C_4(\theta,\phi)   =
\begin{bmatrix}
mrl\sin{(\theta-\Psi)}\dot\Psi+mr^2\sin{(\theta+\phi-\Psi)}\dot\Psi\\
mr^2\sin{(\theta+\phi-\Psi)}\dot\Psi
\end{bmatrix}\notag\\
%&C_{21}=mrl\sin{(\theta_1-\Psi)}\dot\Psi+mr^2\sin{(\theta_1+\phi_1-\Psi)}\dot\Psi\notag\\
%&C_{31}=mr^2\sin{(\theta_1+\phi_1-\Psi)}\dot\Psi\notag\\
%&C_{41}=mrl\sin{(\theta_2-\Psi)}\dot\Psi+mr^2\sin{(\theta_2+\phi_2-\Psi)}\dot\Psi\notag\\
%&C_{51}=mr^2\sin{(\theta_2+\phi_2-\Psi)}\dot\Psi\notag
&M_{11}=I_G+2mr^2.\notag
\end{align}\par
In above equations, $r$ and $l$ denote the radius of the satellite and tether length, respectively. The tether connecting point's moment of inertia is known as $I_r=I_G+mr^2$. \begin{color}{red}Consider matrix determinant of $M_1$
\begin{align*}
\vert M_1\vert
&= (I_r+ml^2+2mrl\cos{\phi})I_r-(I_r+mrl\cos{\phi})^2\\
&= I_rml^2 - m^2r^2l^2\cos^2{\phi}.
\end{align*}
Hence $\vert M_1\vert = I_gml^2+m^2r^2l^2(1-\cos^2{\phi})>0$, and $M_1$ is invertible~\cite{horn2012matrix}. In the view of Lagrangian mechanics, one can get the similar result, since $M_{1}$ is a positive definite matrix~\cite{ortega2013passivity}.\end{color}
\subsection{Model Reduction and Individual System}
Using a simple exponential stabilization controller represses the behaviors of $\Psi$ ($\Psi\approx\theta_i$, $\ddot\Psi\rightarrow 0$) in Eq.~(\ref{eq:3bodymodel}), and the reduced dynamic equations may be derived as~\cite{chung2007nonlinear,huang2015nonlinear}:
\begin{align}
M_1(\phi_i)
\begin{pmatrix}
\ddot{\theta}_i\\
\ddot{\phi}_i
\end{pmatrix}
+C_1(\phi_i,\dot{\theta}_i,\dot{\phi}_i)
\begin{pmatrix}
\dot{\theta}_i\\
\dot{\phi}_i
\end{pmatrix}
=
\begin{pmatrix}
\tau_{\theta i}\\
\tau_{\phi i}
\end{pmatrix}.\label{eq:3bodyreduce}
\end{align}\par
Figure~\ref{Figure:modelsinglebody} shows an individual participating satellite which rotates about the center of the array frame. \begin{color}{red}Body coordinate system $x$-$y$ requires that y axis points to the direction of the increase of the array rotation angle, and meanwhile $x$-axis is along the tether\end{color}. Assume that the tether is massless and inextensible so that the vibration dynamics can be regardless. The Kevlar physical property makes the tether strong and thin enough, i.e., SPECS owns the tether consisting of four tethers lines, and the total mass is approximate to 30 \textit{kg} for 1-km-long ribbon tether\cite{chung2007nonlinear1}. \begin{color}{red}The three-satellite inline tethered frame is assumed to rotate at a specific angular rate so that the tether keeps taut.\end{color} Since the tether speed can be easily controlled by programmed tether motor and doesn't disturb any other system states, it's assumed that the tether length speed $\dot l$ is naught. Finally, assume the three-satellite inline tethered frame is manoeuvred to orbit in the second Lagrangian point of the earth-sun system referring to the SPECS project, and hence the gravity term is approximate to zero~\cite{lorenzini2006far}.\par
%figure2
\begin{figure}
\centering
\includegraphics[width=0.75\textwidth]{paper2_Fig2.eps}
\caption{Schematic representation of three-satellite array dynamics}
\label{Figure:model3body}
\end{figure}
%figure3
\begin{figure}
\centering
\includegraphics[width=0.75\textwidth]{paper2_Fig3.eps}
\caption{Schematic representation of single satellite dynamics}
\label{Figure:modelsinglebody}
\end{figure}
To our knowledge, flywheels are applied as the proper actuators frequently in the attitude regulation of the satellite, for instance, the system of the SPECS project is manoeuvred by using the flywheel control strategy in most of the time~\cite{chung2008propellant1}. In the view of the mechanical structure, there are two categories of flywheels, namely, the bias moment and the zero moment. Functionally, the rotation rate of the bias moment wheels is nonnegative; the zeros moment ones can change the rotation direction. The zero moment wheels are typically known as reaction wheels, which may always maintain the whole satellite's moment zero, and are used more widely than the bias ones in the attitude regulation of the satellite. Considering the energy consumption of the system in practice, \begin{color}{red}flywheels are driven by the electricity which may be provided by energy harvesting system\end{color}, and in contrast with the thruster, the utilization of flywheels avoids consuming the limited quantities of fuel during the attitude regulation. In the aspect of regulating accuracy, the TSS need reaction wheels to generate the continuous high precise torque for the regulation in orbit, whereas thrusters consume expensive fuel to supply the linear torque output in pulse method. In addition, the reaction wheel is also convenient to equip, \begin{color}{red}and there always exists an incline reaction wheel equipped as the redundancy actuator.\end{color} For the attitude regulation, with neglecting the thruster force input, the reaction wheel control strategy changes the three-satellite inline frame into an underactuated system. More notably, the effect of the input for the control system is saturated, since the torque of an reaction wheel is not infinite. Normally, for a reaction wheel, the positive and negative torques are identical, which can be expressed by the saturation function.
\begin{align}
\bar{sat}(u)=
\begin{cases}
T_p &u > T_p\\
u   &T_n\le u\le T_p\\
T_n &u < T_n
\end{cases}
\end{align}
where $T_p$ and $T_n$ denote the maximal positive and negative torque, respectively. According to aforementioned analysis, the equations for the dynamics of the individual satellite shown in Figure~\ref{Figure:modelsinglebody} becomes
\begin{align}
M_1(\phi)
\begin{pmatrix}
\ddot{\theta}\\
\ddot{\phi}
\end{pmatrix}
+C_1(\phi,\dot{\theta},\dot{\phi})
\begin{pmatrix}
\dot{\theta}\\
\dot{\phi}
\end{pmatrix}
=
\begin{bmatrix}
1\\
1
\end{bmatrix}\bar{sat}(u)\label{eq:individualsys}
\end{align}
where $\bar{sat}()$ means the saturation characteristics. Attend to the problem of the reduced three-satellite inline tethered formation, and the following lemma is proposed to demonstrate the stability of the coupled attitude dynamics shown in Eq.~(\ref{eq:3bodymodel}).
\begin{lemma}
Synchronization and Partial Contraction~\cite{wang2005partial}. \begin{color}{red}Consider two nominal coupled systems which have the relation\end{color}
\begin{align}
\dot{x}_1-f(x_1,t)=\dot{x}_2-f(x_2,t)\notag
\end{align}
where the function $f(x,t)$ has contracting property in an input-independent matrix, then the state $x_1$ and $x_2$ will approach each other in spite of the different initial requirement.\label{theorem:syn}
\end{lemma}\par
According to Lemma~\ref{theorem:syn}, Eq.~(\ref{eq:3bodyreduce}) indicates that the formulation of two sub-systems meets the contracting requirement, hence $(\theta_1,\phi_1)^T$ and $(\theta_1,\phi_2)^T$ will coverage to each other if an individual control law is designed to guarantee the stability of the each one~\cite{huang2015nonlinear}.\par
\begin{color}{red}In this work, the linear force is ignored, and hence the system becomes underactuated and input saturated.\end{color}
As $M_1(\phi)$ is invertible, select state variable $x_1=\theta$, $x_2=\dot\theta$, $x_3=\phi$, $x_4=\dot\phi$ to get the following state space equations, which express the attitude dynamics of the individual participating satellite shown in Figure~\ref{Figure:modelsinglebody}:
\begin{align}
\begin{split}
&\dot x_1 = x_2\\
&\dot x_2 = f_1(X,t)+d_1(X,t)+b_1(X,t)\bar{sat}(u)\\
&\dot x_3 = x_4\\
&\dot x_4 = f_2(X,t)+d_2(X,t)+b_2(X,t)\bar{sat}(u)
\end{split}\label{eq:model}
\end{align}
where $X = (x_1,x_2,x_3,x_4)^T$, and $f_1(X,t)$ and $f_2(X,t)$ are nonlinear functions:
\begin{align*}
f_1(X,t) =& \frac{I_rrl(x_2+x_4)\sin x_3+mr^2l^2x_2\cos x_3\sin x_3}{I_rl^2-mr^2l^2\cos{^2x_3}}x_2\\
          &+\frac{I_rx_4 mrl(x_2+x_4)\sin x_3 }{I_rml^2-m^2r^2l^2\cos{^2x_3}}\\
f_2(X,t) =& \frac{(I_r+mrl\cos x_3)mrlx_4\sin x_3-I_r(I_r+ml^2+2mrl\cos x_3)}{m^2r^2l^2\cos{^2x_3}-I_rml^2}x_2\\
          &+\frac{[mrl(x_2+x_4)\sin x_3](I_r+mrl\cos x_3)}{m^2r^2l^2\cos{^2x_3}-I_rml^2}x_4\\
b_1(X,t) =&-\frac{mrl\cos x_3}{I_rml^2-m^2r^2l^2\cos{^2x_3}}\\
b_2(X,t) =&\frac{mrl\cos x_3+ml^2}{I_rml^2-m^2r^2l^2\cos{^2x_3}}
\end{align*}\par
In addition, $d_1(X,t)$ and $d_2(X,t)$ denote the external perturbations such as the solar pressure torque, the aerodynamic torque or the magnetic torque, and among the previous perturbations, the aerodynamic torque plays a major role, which will affect the stability of the attitude.\par
\begin{color}{red}At this point, the attitude regulation of the individual satellite expressed in Eq.~(\ref{eq:model}) has been described as an issue about the stabilization of an underactuated and coupling input system with input saturation.\end{color} Once the states of Eq.~(\ref{eq:model}) is stable, according to Lemma~\ref{theorem:syn}, the entire three-satellite array can be adjusted to be a fixed-point rotation system.
\section{Modified Hierarchical Sliding Mode Control with Input Saturation}\label{sec:SMC}
\begin{color}{red}The case about attitude regulation of the three-satellite inline tethered system is converted into designing the independent control system for the individual satellite described as Eq.~(\ref{eq:model}), and meanwhile, the system is strictly saturated, coupled and nonlinear.\end{color} There exists a stable \begin{color}{red}hierarchical sliding mode (HSM)\end{color} approach investigated for a class of second order underactuated systems, like Acrobot, overhead crane and pendubot etc~\cite{wang2004design}, and the state space equations of the these models are similar to the form of Eq.~(\ref{eq:model}). Considering the effect of the input saturation, synthesize the HSM and adaptive sliding mode control schemes to obtain a modified method, which can overcome both the input saturation and coupling characteristics.\par
In the symbol system of the control system, the time variable is usually abbreviated if it won't lead to confusing, i.e., $S(t)$ may be simplified as $S$. HSM structures the first-level sliding surfaces as follows:
\begin{align}
\begin{split}
s_1 = c_1x_1+x_2\\
s_2 = c_2x_3+x_4,
\end{split}\label{eq:s1s2}
\end{align}
\begin{color}{red}where $c_1$ and $c_2$ are suitable positive parameters,\end{color} and the second-level sliding surface is designed as
\begin{align}
S=\alpha s_1+\beta s_2\label{eq:S},
\end{align}\par
The proceeding of the HSM design is omitted~\cite{wang2004design}. There is an important lemma for the modified controller.
\begin{lemma}Hierarchical Convergence~\cite{wang2004design}.
Considering a sort of underactuated systems which own a stabel equilibrium point, propose the sliding surfaces as Eq.~(\ref{eq:s1s2}) and Eq.~(\ref{eq:S}), and there is a approach to guaranteeing the asymptotically convergence of the second-level sliding mode. If $s_1\in L_\infty$, $\dot s_1\in L_\infty$, the first-level sliding mode also asymptotically converges.\label{lm:1}
\end{lemma}\par
Consider the nominal nonlinear system Eq.~(\ref{eq:model}), and rewrite $\bar{sat}(u)$ as $\bar{sat}(u) = R(u)u$. Note that $0 < R(u) \le 1$
\begin{equation}
R(u) =\begin{cases}
u_{max}/u   & u > u_{max}\\
1           & u_{min} \le u \le u_{max}\\
u_{min}/u   & u < u_{min}
\end{cases}
\end{equation}
where $u_{max}$ $(u_{min})$ is the maximum torque that the reaction flywheel can generate in positive (negative) direction, and there is a restriction $u_{min}<0<u_{max}$, $u_{min}=-u_{max}$ for guaranteeing the predefined condition $0<R(u)\le 1$.\par
There exist the following assumptions in this paper.
\begin{assumption}
The uncertain dynamic system with nonlinearities indicated in Eq.~(\ref{eq:model}) is controllable.\label{as:1}
\end{assumption}
\begin{assumption}
$d_1(X,t)$ and $d_2(X,t)$ are the uncertain parts of system~(\ref{eq:model}), and satisfies the following relation $max(\vert d_1(X,t)\vert,\vert d_2(X,t)\vert) \le \zeta <\infty$, \begin{color}{red}where $\zeta$ is constant positive.\end{color}\label{as:2}
\end{assumption}
\begin{remark}
The mathematical form of $d_i(X,t)$ is trigonometric term, and the detail of $d_i(X,t)$ will be given in Section~\ref{sec:sm}. Assumption~\ref{as:2} is reasonable since there exists the boundary of the perturbation mentioned above. In addition, it is obvious that the terms in the individual satellite system~(\ref{eq:3bodyreduce}) consist of the trigonometric and constant terms, and hence the extremes of $\vert c_1x_2+f_1(X,t)\vert$ and $\vert c_2x_4+f_2(X,t)\vert$ can be obtained by using mathematical calculation.
\end{remark}\par
Modified hierarchical sliding mode control (MHSM) constructs the first- and second-level surfaces like Eq.~(\ref{eq:s1s2}) and Eq.~(\ref{eq:S}), and thanks to the Lemma~\ref{lm:1} and Assumption~\ref{as:2}, we propose the adaptive control input as follows:
\begin{align}
u = -\frac{\rho\delta\eta(t)}{\alpha b_1(X,t)+\beta b_2(X,t)}sgn(S)\label{eq:u}.
\end{align}\par
\begin{color}{red}
$\eta(t)$ is designed as the system gain with the adaptive update rate, and the stability of the system~(\ref{eq:model}) can be guaranteed with its help even if the input saturation occurs during the attitude regulation.
\end{color}
In the preceding expression
\begin{align}
\rho &= 2(\alpha+\beta)max(\zeta,\vert c_1x_2+f_1(X,t)\vert,\vert c_2x_4+f_2(X,t)\vert)
\end{align}
and $\delta>1$. The derivative of $\eta(t)$ has the characteristics $$\dot{{\eta}}(t) = \rho\gamma\delta e^{-2\gamma^{-1}{\eta}^{-1}(t)}\vert S\vert$$ and $\eta(0) > 0$ so that the $\eta(t)$ is constant positive. $\gamma$ is a parameter for cancelling the input saturation, and it will be discussed later on. \begin{color}{red}
For voiding the singularity of the input~(\ref{eq:u}), the selection of $\alpha$ and $\beta$ may satisfy the condition: $0<\alpha\le\beta$ with $\vert x_3\vert<\pi/2$.
    Selecting $\alpha$ and $\beta$ with $0<\alpha\le\beta$ yields
    \begin{align*}
    &\alpha b_1(X,t)+\beta b_2(X,t)\\
    =& \alpha(b_1(X,t)+b_2(X,t))+(\beta-\alpha)b_2(X,t)\\
    =&\frac{(\beta-\alpha)mrl\cos x_3 + \beta ml^2}{I_rml^2-m^2r^2l^2\cos{^2x_3}}>0.
    \end{align*}
    This can guarantee the non-singularity of Eq.~(\ref{eq:u}), when $x_3$ is varying in a small region with $\vert x_3\vert<\pi/2$. For avoiding the singularity absolutely, one can select $\alpha = \beta >0$ to remove the effect of $x_3$.
\end{color}
\begin{theorem}
The nonlinear system which can be expressed as Eq.~(\ref{eq:model}), satisfies the Assumption~\ref{as:1} and~\ref{as:2}, and if the candidate of system input is Eq.~(\ref{eq:u}), the trajectories of the system asymptotically converge to the sliding surface $S = 0$.
\end{theorem}
\begin{proof}
Select the Lyapunov function candidate
\begin{equation}
V = \frac{1}{2}S^2+\frac{1}{2}v^2\label{eq:V}
\end{equation}
where $v = {\eta}(t)e^{\gamma^{-1}{\eta}^{-1}(t)}$, then introduce Eq.~(\ref{eq:S}) and Eq.~(\ref{eq:u}) into the derivative of $V$ to obtain
\begin{align}
\begin{split}
\dot{V} &= S\dot S+v\dot v\\
        &= S[\alpha(c_1x_2+f_1(X,t)+d_1(X,t)+b_1(X,t)\bar{sat}(u))\\
        &\quad+\beta(c_2x_3+f_2(X,t)+d_2(X,t)+b_2(X,t)\bar{sat}(u))]+v\dot v\\
        &\le \rho\vert S\vert - R(u)\rho\delta\eta(t)\vert S\vert +v\dot v\label{eq:dotV}
\end{split}
\end{align}\par
It is explicit to realize that $0 < R(u(t)) \le 1$, and a constant positive $\gamma$ can be selected to satisfy
\begin{equation}
0<\gamma\le R(u)\le 1 \label{eq:utl}
\end{equation}\par
Moreover, term $R(u)\rho\delta\eta(t)\vert S\vert$ is positive, and $0< R(u)\le 1$. \begin{color}{red}The inequality.~(\ref{eq:dotV}) becomes the interim form:\end{color}
\begin{align}
\dot{V} \le \rho\vert S\vert - \rho\gamma\delta\eta(t)\vert S\vert +v\dot v\label{eq:dotVinterim}
\end{align}\par
Derive $v\dot v$ as follows
\begin{align}
\begin{split}
v\dot v &= \dot{\eta}e^{2\gamma^{-1}\eta^{-1}(t)}(\eta(t)-\gamma^{-1})\\
        &= \rho\gamma\delta(\eta(t)-\gamma^{-1})\vert S\vert\label{eq:vdotv}
\end{split}
\end{align}\par
Substituting Eq.~(\ref{eq:vdotv}) into Eq.~(\ref{eq:dotVinterim}) yields
\begin{align}
\dot{V} \le (1 -\delta)\rho\vert S\vert\le 0 \label{eq:dotVfinal}
\end{align}
$\dot{V}$ is negative semi-definite while $V$ is positive definite, and moreover, bounded functions $S$ and ${\eta}(t)$ ensure that $\lim\limits_{t\to\infty}V$ exists. Inequality.~(\ref{eq:dotVfinal}) can be integrated
\begin{align}
\begin{split}
V(0) &\ge V(t) + \int_{0}^{t}(\delta-1)\rho\vert S(\tau)\vert d\tau\\
&\ge \int_{0}^{t}(\delta-1)\rho\vert S(\tau)\vert d\tau\label{eq:V10}
\end{split}
\end{align}\par
According to the description of Barbalat's lemmas, $t\to\infty$ implies $S(t)\to 0$. Hence, proof is demonstrated completely.
\end{proof}
\section{Adaptive Sliding Mode Control For Two-satellite Array System}\label{sec:SMC2}
The three-satellite inline array can be simplified as three individual satellite systems by above-mentioned model reduction technology, and in this section, a two-satellite array system shown in Figure~\ref{Figure:model2body} is investigated. The two satellites are the participants connected with a stiff space tether, and the dynamic equations can be described as follows:
%figure4
\begin{figure}
\centering
\includegraphics[width=0.75\textwidth]{paper2_Fig4.eps}
\caption{Schematic representation of two-satellite system dynamics}
\label{Figure:model2body}
\end{figure}
\begin{align}
M_3(\phi_1,\phi_2)
\begin{pmatrix}
\ddot{\theta}\\
\ddot{\phi}_1\\
\ddot{\phi}_2
\end{pmatrix}
&+C_5(\phi_1,\phi_2,\dot{\theta},\dot{\phi}_1,\dot{\phi}_2)
\begin{pmatrix}
\dot{\theta}\\
\dot{\phi}_1\\
\dot{\phi}_2
\end{pmatrix}\notag\\
&+
\begin{pmatrix}
2m(r\cos\phi_1+r\cos\phi_2+2l)\dot\theta\dot l\\
2mr\cos\phi_1\dot\theta\dot l\\
2mr\cos\phi_2\dot\theta\dot l
\end{pmatrix}
=
\begin{pmatrix}
\tau_{\phi_1}+\tau_{\phi_2}\\
\tau_{\phi_1}\\
\tau_{\phi_2}
\end{pmatrix}\label{eq:2body}
\end{align}
where
\begin{align}
&M_3(\phi_1,\phi_2)
=\begin{bmatrix}
m_{11}(\phi_1)+m_{11}(\phi_2) &m_{12}(\phi_1) &m_{12}(\phi_2)\\
m_{21}(\phi_1) &m_{22} &0\\
m_{21}(\phi_2) &0      &m_{22}
\end{bmatrix}\label{eq:M3}\\
&C_5(\phi_1,\phi_2,\dot{\theta},\dot{\phi}_1,\dot{\phi}_2)
=\begin{bmatrix}
c_{11}(\phi_1,\dot{\phi}_1)+c_{11}(\phi_2,\dot{\phi}_2) &c_{12}(\phi_1,\dot{\theta},\dot{\phi}_1) &c_{12}(\phi_2,\dot{\theta},\dot{\phi}_2)\\
c_{21}(\phi_1,\dot{\theta}) &c_{22} &0\\
c_{21}(\phi_2,\dot{\theta}) &0 &c_{22}
\end{bmatrix}\label{eq:C5}
\end{align}
and the elements $m_{11}=m_{12}$, $m_{22}$, $m_{21}$, $c_{11}$, $c_{12}$, $c_{21}$ and $c_{22}$ in Eq.~(\ref{eq:M3}) and Eq.~(\ref{eq:C5}) are defined in the attitude dynamics of the individual satellite in Eq.~(\ref{eq:M1}) and Eq.~(\ref{eq:C1}).\par
Compared to the issue about stabilization of the three-satellite inline array, the central satellite is absent in the two-satellite system so that the individual torque acting on one participant will effect the other one~\cite{chung2007nonlinear}. With model reduction techniques failing in this model, the systemical simplification is unable to be achieved due to the fact that the physical construction of the satellite is almost identical. \begin{color}{red}The same mechanical structure of the satellite results in that the system do not satisfy the condition of station keeping like the simple mother satellite and sub-satellite tethered system.\end{color} In the mother satellite and sub-satellite tethered system, it requires the mass of mother satellite to be much larger than the other one, and this premise of assumption is usually met in the deployment mission of the tethered satellite system. Without any simplification, the number of the input in Eq.~(\ref{eq:2body}) is less than the degree of freedom (DOF), and meanwhile the input is coupled as well. For the reason of generality, \begin{color}{red}the condition of the actuator is similar to that in the three-satellite inline array. On the basis of the previous results, the two-satellite inline system is the underactuated system with input saturation.\end{color}\par
The solution for the attitude stability of the three-satellite inline array is valid for the two-satellite inline system. For the sake of the adaptive controller design, the system can be rewritten as the expression of Eq.~(\ref{eq:model}). Select state variables $x_1=\theta$, $x_2=\dot\theta$, $x_3=\phi_1$, $x_4=\dot\phi_1$, $x_5=\phi_2$, $x_6=\dot\phi_2$ to get the following state space equations:
\begin{align}
\begin{split}
&\dot x_1 = x_2\\
&\dot x_2 = f_1(X,t)+d_1(X,t)+b_1(X,t)\bar{sat}(u_1)+b_2(X,t)\bar{sat}(u_2)\\
&\dot x_3 = x_4\\
&\dot x_4 = f_2(X,t)+d_2(X,t)+b_3(X,t)\bar{sat}(u_1)+b_4(X,t)\bar{sat}(u_2)\\
&\dot x_5 = x_6\\
&\dot x_6 = f_3(X,t)+d_3(X,t)+b_5(X,t)\bar{sat}(u_1)+b_6(X,t)\bar{sat}(u_2)
\end{split}\label{eq:2model}
\end{align}
where $X = (x_1,x_2,x_3,x_4,x_5,x_6)^T$. \begin{color}{red}In addition, $\bar{sat}()$, $f_i(X,t)$ and $d_i(X,t)$ have the similar mathematical featureS which are illustrated in the individual system of three-satellite inline array in Eq.~(\ref{eq:model}),\end{color} and $f_i(X,t)$ and $b_i(X,t)$ can be derived:
\begin{align}
\begin{split}
\begin{pmatrix}
f_1(X,t)\\
f_2(X,t)\\
f_3(X,t)
\end{pmatrix}=
M_3^{-1}(x_3,x_5)C_5(x_3,x_5,x_2,x_4,x_6)
\begin{pmatrix}
x_2\\
x_4\\
x_6
\end{pmatrix}
\end{split}
\end{align}
\begin{align}
\begin{split}
\begin{pmatrix}
b_1(X,t)&b_2(X,t)\\
b_3(X,t)&b_4(X,t)\\
b_5(X,t)&b_6(X,t)
\end{pmatrix}=
M_3^{-1}(x_3,x_5)
\begin{pmatrix}
1 &1\\
1 &0\\
0 &1
\end{pmatrix}
\end{split}
\end{align}
where $M_3^{-1}(x_3,x_5)$ is the inverse matrix of $M_3(x_3,x_5)$:
\begin{footnotesize}
\begin{align*}
&M_3^{-1}(x_3,x_5)
=\frac{\begin{bmatrix}
m_{22}^2 &-m_{12}(x_3)m_{22} &m_{22}m_{12}(x_5)\\
-m_{22}m_{21}(x_3) &[m_{11}(x_3)+m_{11}(x_5)]m_{22}-m_{12}^2(x_5) &m_{12}(x_5)m_{21}(x_3)\\
-m_{22}m_{21}(x_5) &m_{12}(x_3)m_{21}(x_5) &[m_{11}(x_3)+m_{11}(x_5)]m_{22}-m_{12}^2(x_3)
\end{bmatrix}}{[m_{11}(x_3)+m_{11}(x_5)]m_{22}^2-m_{12}(x_5)m_{21}(x_5)m_{22}-m_{12}(x_3)m_{21}(x_3)m_{22}}\\
&=\frac{\begin{bmatrix}
I_r^2 &-I_r(I_r+mrl\cos\phi_1) &-I_r(I_r+mrl\cos\phi_2)\\
-I_r(I_r+mrl\cos\phi_1) &I_r(I_r+2ml^2+2mrl\cos\phi_1)-m^2r^2l^2\cos^2\phi_2 &(I_r+mrl\cos\phi_2)(I_r+mrl\cos\phi_1)\\
-I_r(I_r+mrl\cos\phi_2) &(I_r+mrl\cos\phi_2)(I_r+mrl\cos\phi_1) &I_r(I_r+2ml^2+2mrl\cos\phi_2)-m^2r^2l^2\cos^2\phi_1
\end{bmatrix}}{I_rml^2(2I_r-mr^2\cos{^2\phi_1}-mr^2\cos{^2\phi_2})}.
\end{align*}
\end{footnotesize}\par
\begin{color}{red}Consider the denominator of $M_3^{-1}(x_3,x_5)$:
\begin{align*}
&I_rml^2(2I_r-mr^2\cos{^2\phi_1}-mr^2\cos{^2\phi_2})\\
=& I_rml^2(2I_g+2mr^2-mr^2\cos{^2\phi_1}-mr^2\cos{^2\phi_2})\\
=& I_rml^2[2I_g+mr^2(1-\cos{^2\phi_1})+mr^2(1-\cos{^2\phi_2})]>0
\end{align*}
Hence $M_{3}$ is invertible.\end{color}\par
For applying the MHSM control to the attitude regulation of the two-satellite inline system shown in Eq.~(\ref{eq:2model}), the first-level iedal sliding surfaces are selected firstly as follows:
\begin{align}
\begin{split}
\sigma_1 = \bar{c}_1x_1+x_2\\
\sigma_2 = \bar{c}_2x_3+x_4\\
\sigma_3 = \bar{c}_3x_4+x_5
\end{split}\label{eq:sigma123}
\end{align}
where $\bar{c}_i$ is the suitable positive coefficient, and the second sliding surface follows:
\begin{align}
\bar{S} = \bar{\alpha}\sigma_1+ \bar{\beta}\sigma_2+\bar{\iota}\sigma_3\label{eq:bar S}
\end{align}
where $\bar{\alpha}$, $\bar{\beta}$ and $\bar{\iota}$ are positive coefficients.
\begin{remark}
Some essential results are given in advance so that the adaptive control design for the system~(\ref{eq:2model}) can be illustrated explicitly. The spatial perturbation term, $d_i(X,t)$, owns the characteristics of trigonometric function, and those bound value can be determined easily. More notably, the terms in the attitude dynamics of the two-satellite system~(\ref{eq:2body}) are a group of constant and trigonometrical terms, and furthermore, their operations are additive and multiplicative. Hence the boundary exists and can be calculated. In the regulation, the attitude angle is limited in a practical meaningful range $\vert \phi_i\vert<\pi/2$~\cite{chung2008propellant2}. Therefore, some symbols are introduced to express the mentioned relationship, and they will take part in the design of the adaptive control in the following form:\label{rem:2}
\begin{align}
\begin{split}
\bar\varpi &\ge \max(\vert d_1(X,t)\vert,\vert d_2(X,t)\vert,\vert d_3(X,t)\vert)\\
\bar\xi &=  \max(\vert \bar c_1x_2 + f_1(X,t)\vert,\vert \bar c_2x_4 + f_2(X,t)\vert,\vert \bar c_3x_6 + f_3(X,t)\vert)
\end{split}
\end{align}
\end{remark}\par
The adaptive control inputs for stabilizing the attitude dynamics in Eq.~(\ref{eq:2model}) are given directly:
\begin{align}
\begin{split}
u_1 &= -\frac{\bar\delta\bar\rho\eta_1(t)sgn(\bar S)+k_1\bar S}{\bar\alpha b_1(X,t) + \bar\beta b_3(X,t) + \bar\iota b_5(X,t)}\\
u_2 &= -\frac{\bar\delta\bar\rho\eta_2(t)sgn(\bar S)+k_2\bar S}{\bar\alpha b_2(X,t) + \bar\beta b_4(X,t) + \bar\iota b_6(X,t)}.
\end{split}\label{eq:u1u2}
\end{align}\par
\begin{color}{red}
For voiding the singularity of the input, the parameter $\bar\alpha$, $\bar\beta$ and $\bar\iota$ should satisfy the the condition $0<\bar\alpha\le\bar\iota$ and $0<\bar\alpha\le\bar\beta$ when $\vert\phi_1\vert<\pi/2$ and $\vert\phi_2\vert<\pi/2$.\par
    Consider the expression $\bar\alpha b_1(X,t) + \bar\beta b_3(X,t)$ with $0<\bar\alpha\le \bar\beta$
    \begin{align*}
    &\bar\alpha b_1(X,t) + \bar\beta b_3(X,t)\\
    =&-\bar\alpha I_rmrl\cos\phi_1 + \bar\beta I_rmrl\cos\phi_1+\bar\beta[ml^I_r+ml^2I_g+m^2r^2l^2(1-\cos^2\phi_2)]\\
    =&(\bar\beta-\bar\alpha) I_rmrl\cos\phi_1+\bar\beta[ml^I_r+ml^2I_g+m^2r^2l^2(1-\cos^2\phi_2)]>0
    \end{align*}
    when $\vert\phi_1\vert<\pi/2$ and $\vert\phi_2\vert<\pi/2$.
    In addition, it is obvious that $b_5(X,t)>0$, and hence, for voiding the singularity of $\bar\alpha b_1(X,t) + \bar\beta b_3(X,t) + \bar\iota b_5(X,t)$,  the parameters should satisfy the condition $0<\bar\alpha\le\bar\beta$ and $\bar\iota>0$. $\bar\alpha b_2(X,t) + \bar\beta b_4(X,t) + \bar\iota b_6(X,t)>0$ can be obtained under the condition: $0<\bar\alpha\le\bar\iota$ and $\bar\beta>0$. Conclusively, when $\vert\phi_1\vert<\pi/2$ and $\vert\phi_2\vert<\pi/2$, $0<\bar\alpha\le\bar\iota$ and $0<\bar\alpha\le\bar\beta$ guarantee the non-singularity of the input in the MHSM control, and if one wants to obtain the non-singularity absolutely, $0<\bar\iota=\bar\alpha=\bar\beta$ is essential.
\end{color}
In preceding equation~(\ref{eq:u1u2})
\begin{align}
\bar\rho &\ge 2(\bar\alpha + \bar\beta + \bar\iota)\max(\bar\varpi,\bar\xi)
\end{align}
and $\bar\eta(t)=\eta_1(t)+\eta_2(t)$ whose derivative has the characteristics, $$\dot{{\bar\eta}}(t) = \bar\rho\bar\gamma\bar\delta e^{-2\bar\gamma^{-1}{\bar\eta}^{-1}(t)}\vert \bar S\vert,$$ $\bar\eta(0) > 0$ so that $\bar\eta(t)$ is a positive constant. $k_1+k_2\ge 0$, $\bar\delta >1$, and in addition, the input saturation can be described by $\bar{sat}(u_i) = R(u_i)u_i, i =1,2$, whose bound of $R(u_i)$ has the characteristics, $0<\bar\gamma\le \min(R(u_1),R(u_2))\le 1$. The crucial theorem follows.
\begin{theorem}
Considerinig the nonlinear system which is shown as Eq.~(\ref{eq:2model}) and the constraint conditions of the system established in Remark~\ref{rem:2}, if the candidate of the system inputs are expressed as Eq.~(\ref{eq:u1u2}), the trajectories of the system will asymptotically converge to the ideal sliding surface $\bar S=0$.\label{theorem:3}
\end{theorem}
\begin{proof}
Select the Lyapunov function candidate
\begin{align}
\bar V = \frac{1}{2}\bar S^2 + \frac{1}{2}\bar v^2
\end{align}
where $\bar v = {\bar\eta}(t)e^{\bar\gamma^{-1}{\bar\eta}^{-1}(t)}$, then introduce Eq.~(\ref{eq:bar S}) and Eq.~(\ref{eq:u1u2}) into the derivative of $V$ to obtain
\begin{align}
\begin{split}
\dot{\bar V} &= \bar{S}\dot{\bar{S}}+\bar v\dot{\bar v}\\
                &= \bar{S}(\bar\alpha[\bar c_1 x_2 + f_1(X,t) + d_1(X,t) + b_1(X,t)\bar{sat}(u_1) + b_2(X,t)\bar{sat}(u_2)]\\
                     &\quad+         \bar\beta [\bar c_2 x_4 + f_2(X,t) + d_2(X,t) + b_3(X,t)\bar{sat}(u_1) + b_4(X,t)\bar{sat}(u_2)]\\
                     &\quad+         \bar\iota[\bar c_3 x_6 + f_3(X,t) + d_3(X,t) + b_5(X,t)\bar{sat}(u_1) + b_6(X,t)\bar{sat}(u_2)])+\bar v\dot{\bar v}\\
                     &\le \bar\rho\vert\bar S\vert + \bar S[(\bar\alpha b_1(X,t) + \bar\beta b_3(X,t) + \bar\iota b_5(X,t))R(u_1)u_1\\
                     &\quad+(\bar\alpha b_2(X,t) + \bar\beta b_4(X,t) + \bar\iota b_6(X,t))R(u_2)u_2]+\bar v\dot{\bar v}\\
                     &\le \bar\rho\vert\bar S\vert - \bar S[ R(u_1)(\bar\delta\bar\rho\eta_1(t)sgn(\bar S)+k_1\bar S)+R(u_2)(\bar\delta\bar\rho\eta_2(t)sgn(\bar S)+k_2\bar S)]+\bar v\dot{\bar v}\\
                     &\le \bar\rho\vert\bar S\vert - \gamma(k_1+k_2)\bar S^2 - \bar\delta\bar\rho\bar\gamma\bar\eta(t)\vert\bar S\vert+\bar v\dot{\bar v}.
\end{split}
\end{align}\par
Referring to Eq.~(\ref{eq:vdotv}), above expression can be derived to
\begin{align}
\dot{\bar V} \le (1 -\bar\delta)\bar\rho\vert\bar S\vert- \bar\gamma(k_1+k_2)\bar S^2\le 0 .\label{eq:dotbarVfinal}
\end{align}\par
According to the description of Barbalat's lemmas and referring to Eq.~(\ref{eq:V10}), $t\to\infty$ implies $S(t)\to 0$. Hence the second-level sliding surface in Eq.~(\ref{eq:bar S}) is stable, and proof is demonstrated completely.
\end{proof}
\begin{remark}
In addition, if the inputs are selected as:
\begin{align}
\begin{split}
u_1 &= -\frac{\bar\delta\bar\rho\eta_1(t)sgn(\bar S)+k_1\bar S_1}{\bar\alpha b_1(X,t) + \bar\beta b_3(X,t) + \bar\iota b_5(X,t)}\\
u_2 &= -\frac{\bar\delta\bar\rho\eta_2(t)sgn(\bar S)+k_2\bar S_2}{\bar\alpha b_2(X,t) + \bar\beta b_4(X,t) + \bar\iota b_6(X,t)}.
\end{split}\label{eq:modified_u1u2}
\end{align}
where $k_1\bar S_1+k_2\bar S_2=k_0\bar S$ with $k_0>0$, one can also obtain the result above.
\end{remark}
\begin{remark}
Theorem~\ref{theorem:3} is proposed for guaranteeing that the sliding mode dynamics will reach the ideal surface $\bar S$ asymptotically, \begin{color}{red}and the adaptive update ${\bar\eta}(t)$ eliminates the effect of the input saturation in the two-satellite system.\end{color} Moreover, considering the convergence of the first-level sliding mode dynamics, and according to the article~\cite{wang2004design}, they are asymptotically stable, and the proof is omitted here.
\end{remark}
\section{Numerical Simulation}\label{sec:sm}
For exhibiting the effectiveness of the proposed control schemes designed for the multi-satellite tethered systems, numerical simulations are implemented in this section, and furthermore the analysis of simulation results verifies the preceding design in Section~\ref{sec:SMC} and Section~\ref{sec:SMC2}.\par
\begin{table}
% table caption is above the table
\caption{SPHERES properties}\label{ta:properites}
\label{tab:1}       % Give a unique label
% For LaTeX tables use
\begin{tabular}{lll}
\hline\noalign{\smallskip}
Properties              &RWA equiped\\
\noalign{\smallskip}\hline\noalign{\smallskip}
Diameter   &0.25 \textit{m}\\
Mass       &15.856 \textit{kg}\\
Moment of inertia        &0.178 \textit{kg}$\cdot m^2$\\
Max torque actuation   &0.207 \textit{Nm}\\
\noalign{\smallskip}\hline
\end{tabular}
\end{table}
Referring to the experiment of SPHERES project, the simulation parameters describing the physical characteristics of the satellite are listed in Table~\ref{tab:1}. According to the parameter list, the input boundary is determined as $T_p = -T_n = 0.207Nm$. In particular the controller's output is programmed to satisfy the restriction of the actuator, which can avoid the block of the flywheels in some extent. The length of the tether is constant, since the variety of the length rate leads to an extremely complicated and unstable multi-satellite tethered systems. The reasonable length of the tether is essential, $l = 0.9m$, although the length will not take an important role in the following process of validating. In this simulation, it is assumed that the satellites expose itself to the real space environment, and there exist the solar radiation pressure torque and aerodynamic torque effecting its attitude stability. \begin{color}{red}Without loss of generality, the perturbation term is thought to be a periodic function, and the period is $5000s$ approximately, which equals to the orbital period~\cite{inamori2015magnetic,liu2013calculation}. Hence for verifying the effectiveness of the MHSM control, the satellite is exerted by a sinusoidal perturbation whose amplitude is no more than $1\times 10^{-5}Nm$ with $2\times 10^{-4}Hz$. Since the relative distance among the participating satellites is not large, without loss of generality, the form of preceding perturbation is identical.
\begin{align}
d_i(X,t) \approx 2\times 10^{-5}\sin(2\pi/5000)\notag
\end{align}
\end{color}
\subsection{Individual System of Three-satellite Inline Array}
\begin{color}{red}In order to apply the MHSM scheme proposed in Section~\ref{sec:SMC},\end{color} the parameters of the controller is exhibited, $c_1 = 10$ and $c_2 = 3$, which meet the demand of Eq.~(\ref{eq:s1s2}), the parameters of second-level sliding surface in Eq.~(\ref{eq:S}), $\alpha=\beta=1$. The adaptive parameters can be decided as follows: $\delta = 1.8$, $\eta(0) = 1e^{-4}$, $\zeta = 1$ and $\gamma = 1\times 10^{-4}$. $\rho$ obtained due to the simulation process, which satisfies the conditions of Lyapunov stability.\par
%figure5
\begin{figure}
\centering
\includegraphics[width=0.9\textwidth]{paper2_Fig5.eps}
\caption{Three-satellite inline array: curves of all sliding planes}
\label{Figure:sm}
\end{figure}
%figure6
\begin{figure}
\centering
\includegraphics[width=0.9\textwidth]{paper2_Fig6.eps}
\caption{Three-satellite inline array: attitude angle and array angular rate}
\label{Figure:attitude}
\end{figure}
%figure7
\begin{figure}
\centering
\includegraphics[width=0.9\textwidth]{paper2_Fig7.eps}
\caption{Three-satellite inline array: control torque of MHSM}
\label{Figure:Torque}
\end{figure}
In this simulation, the three-satellite inline tethered system maintains rotating at angular rate $\omega_0=0.25$ in initial stage, and meanwhile, the initial individual angle, $\phi_0 = 0.2$, means that the individual system has been swung to \begin{color}{red}an\end{color} undesired attitude. The MHSM scheme aims to regulate the system to the specified attitude, and in particular the desired conditions are as follows, $\omega_d = 0.25$, $\phi_d = 0$ and $\dot{\phi}_d = 0$. In final stage, the entire system will keep rotating, and meanwhile three individual satellites are nonskew and in a line.\par
Use MHSM scheme to regulate the individual system shown as Eq.~(\ref{eq:individualsys}), and the guidance situation of sliding mode is illustrated in Figure~\ref{Figure:sm} where $s_1$ and $s_2$ converge to origins as well as $S$. It is the trajectory of $s_1$ that nearly overlaps with the one of $S$, and $s_2$ fluctuates slightly in the vicinity of the origin. Paying attention on the convergence of each sliding mode, it is obvious that $S$, $s_1$ and $s_2$ reach the neighbour region of the origin at about 1.2 $s$ and stay stable. Incidentally, the state of the angle and the torque are both physical meaningful.\par
The situation of the attitude regulation is illustrated in Figure~\ref{Figure:attitude}, in which the array angular rate is adjusted to be the desired value from the initial condition so that the individual satellites and the central satellite are rotating together at the same velocity. It is at about 1.2 $s$ when the trajectory of the attitude angle converges to the preset value, $\phi_d = 0$, so does the array angular rate. \begin{color}{red}The performance of the decentralized control reported in~\cite{chung2008propellant1} is also illustrated in Figure~\ref{Figure:attitude}. In~\cite{chung2008propellant1}, a pair of thrusters are used to dump the angular momentum caused by the saturation of the flywheel speed, and for reducing the consumption of thruster firing, a little gain design is demanded in the decentralized control strategy. Hence, the decentralized control inevitably leads to a slow smooth attitude regulation shown in Figure~\ref{Figure:attitude}. Compared with the the decentralized control scheme, one can find that the MHSM control demonstrates a fast regulation, and although the amplitudes of the varying attitude angle and the array angular rate of the MHSM control are greater with the variations falling in $[-0.05,0.2]$ $rad$ and $[0.225,0.375]$ $rad/s$, respectively, this variations are tolerable in some fast attitude scenarios of the TSS~\cite{Williams2009745}. Even if the input saturation occurs during the attitude regulation, the MHSM control scheme guarantees the stability of the system, and hence the selection of the gain is flexible so that the performance of the control system can be adjusted by the suitable parameters. \end{color}\par
The characteristics of the input saturation are exhibited in Figure~\ref{Figure:Torque}, and the existence of the input bound makes the curves not smooth any more. In particular, \begin{color}{red}it cuts off the oversize input signal to be a bound value, and the input resumes to be regular once the control signal is located in the region $[T_n,T_p]$.\end{color} The boundary of the input is decided by the physical property of the flywheel equipped in the satellite, and the bound value is $0.207Nm$ bidirectionally in this section. In Figure~\ref{Figure:Torque}, the saturation occurs from about $0s$ to $0.6s$, which is zoomed out for examining more details, and the torque keeps a low-amplitude fluctuation in the neighbour region of the origin after 1.2 $s$. That means that the torque utilized to make the system stable, still performs accurately, although the stability of the attitude dynamics for the individual satellite has achieved macroscopically. It is also easy to find that the perturbations don't introduce distinct contaminations into the control system, in another word, the anti-interference ability of the proposed scheme is excellent.
\subsection{Two-satellite System}
The parameter set of the MHSM scheme is given firstly for stabilizing the attitude of the two-satellite system expressed in Eq.~(\ref{eq:2model}) in Section~~\ref{sec:SMC2}. The set of the parameters for first-level sliding surface design, $\bar c_1 = 5$, $\bar c_2 =3$ and $\bar c_3=1$, satisfies the positive requirement, and meanwhile, for the the second-level sliding surface design $\bar\alpha$$=\bar\beta=\bar\iota=1$ will not effect the performance of the control system in some extent. List adaptive parameters as follows: $\bar\delta = 2.1$, $\bar\eta(0)=5\times 10^{-4}$, $\bar\gamma = 1\times 10^{-4}$ and $\bar\varpi = 2$. In addition, $\bar\xi,\bar\rho$ can be obtained due to the simulation process. All mentioned parameters guarantee Lyapunov stability of the system dynamics.\par
%figure8
\begin{figure}
\centering
\includegraphics[width=0.9\textwidth]{paper2_Fig8.eps}
\caption{Two-satellite system: attitude angle and array angular rate}
\label{Figure:2attitude}
\end{figure}
%figure9
\begin{figure}
\centering
\includegraphics[width=0.9\textwidth]{paper2_Fig9.eps}
\caption{Two-satellite system: curves of all sliding planes}
\label{Figure:2s}
\end{figure}
%figure10
\begin{figure}
\centering
\includegraphics[width=0.9\textwidth]{paper2_Fig10.eps}
\caption{Two-satellite system: control torques of MHSM}
\label{Figure:2u1u2}
\end{figure}
In this part, the system is fixed-point rotating at the desired angular rate $\dot\theta_d = 0.25$, while the tether keeps stiff. In the final stage, $\phi_{i}$, the attitude angle, should be regulated to be 0.  Initially, the system rotates at the angular rate $\dot\theta_0=0.25$, and the attitude can be described by $\phi_{10}=0.4$ and $\phi_{20}=0.1$.\par
The performance of the attitude regulation can be seen from Figure~\ref{Figure:2attitude}. \begin{color}{red}It shows that the angular rate of the entire system is stable at 1.2 $s$, and meanwhile, the attitude of the participating satellites is adjusted to the desired status. Compared with the decentralized control strategy, similar to the case of the three-satellite system, the MHSM control performs the fast regulation with reasonable amplitudes of attitude angle and array angular rate.\end{color} Figure~\ref{Figure:2s} illustrates the guidance situation of the sliding mode, which indicates that $\sigma_i,i=1,2,3$ and $\bar S$ all converge as desired, and then stay in the neighbour region of the origin in the subsequent time. It is noted that the first-level sliding surfaces are stable in the same time, and obviously this phenomenon proceeds to verify Theorem~\ref{theorem:3}. Figure~\ref{Figure:2u1u2} exhibits the torques of the flywheels equipped in the participating satellites. According to the saturation characteristics of actuators, there exists the bounds of the inputs in the positive and negative direction, and the bounded inputs can be found at about $0.1s$ in this figure.
\section{Conclusion}\label{sec:Conclusion}
This paper proposes a torque control scheme to adjust the attitude of the multi-satellite array system. In particular, the proposed approach accounts for two meaningful issues in practical implementation. The first issue is that the attitude dynamics of the multi-satellite array system is coupled and underactuated during the regulation process. The second one is that the actuators are usually saturated or bounded in the realistic application. For solving underactuated problems, the adaptive rate and the HSM control law are synthesized to obtain the MHSM control scheme, and it also takes into consideration the input saturation. The adaptive rate term is designed to compensate the impact of the saturated input so that the Lyapunov stability of the attitude dynamics can be obtained. Theoretical analysis is demonstrated to guarantee that the sliding mode dynamics reaches the ideal sliding surface. The simulation results indicate that the attitude regulation for the tethered satellite system with the control scheme effectively and efficiently.
%Theoretical analysis is exhibited to guarantee that the states of inline tethered system reach the desired sliding mode in finite time and stay stability. The simulation results indicate that the regulating mission with the control scheme effectively and efficiently.
%The MHSM  control laws have been proposed for the attitude adjusting of multi-satellite tethered system in this paper. The procedure of control design considering  Theoretical analysis is exhibited to guarantee that the states of inline tethered system reach the desired sliding mode in finite time and stay stability. The simulation results indicate that the regulating mission with the control scheme effectively and efficiently. In addition,
%\section{Acknowledgment}
%This work is partially supported by the National Natural Science Foundation of China (No. 61104112, 61503097).
%
% For tables use

\begin{acknowledgements}
%If you'd like to thank anyone, place your comments here
%and remove the percent signs.
This work is partially supported by the National Natural Science Foundation of China (No. 61104112, 61503097).
\end{acknowledgements}
\begin{color}{red}
% BibTeX users please use one of
%\bibliographystyle{spbasic}      % basic style, author-year citations
\bibliographystyle{spmpsci1}      % mathematics and physical sciences
%\bibliographystyle{spmpsci2}
%\bibliographystyle{spphys}       % APS-like style for physics
%\bibliography{}   % name your BibTeX data base

% Non-BibTeX users please use
%\begin{thebibliography}{}
%\bibliographystyle{spbasic}
%\bibliographystyle{spmpsci}
\bibliography{ref_light}
\end{color}
%
% and use \bibitem to create references. Consult the Instructions
% for authors for reference list style.
%
%\bibitem{RefJ}
% Format for Journal Reference
%Author, Article title, Journal, Volume, page numbers (year)
% Format for books
%\bibitem{RefB}
%Author, Book title, page numbers. Publisher, place (year)
% etc
%\end{thebibliography}

\end{document}
% end of file template.tex